---
title: "ESM232 - HwK5"
author: "Ruben Sanchez, Alex Millward and Mauricio Collado"
date: "4/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

# load libraries

library(tidyverse)
library(purrr)
library(deSolve)
```

### 1. Objective

Implement a differential equation for growth rate incorporating carrying capacity. Then predict the population for the next 50 years employing the following set of parameters:<br>

- Initial population of 1
- Intrinsic growth rate of 0.05
- Carrying capacity of 20<br>


### 2. Change in population
<br>
We create a growth rate function with initial population (P0), intrinsic growth (r), and carrying capacity (k) as parameters.<br>

```{r}
#'  Population growth rate
#' @param T  period of growth
#' @param P initial population
#' @param parms$r - base growth rate
#' @param parms$k - carrying capacity
#' @return change in population 
#'
dexpopk = function(time, P, parms) {
  
  # compute rate of change of population that includes carrying capacity
  dexpop = parms$r*P*(1-P/parms$k)
  
  return(list(dexpop))
}

```

### 3. Prediction for 50 years
<br>
We predict the population for next 50 years with the following parameters:<br>

- Initial population of 1
- Intrinsic growth rate of 0.05
- Carrying capacity of 20

```{r}
# set parameters
initial_pop=1
params_rk = list(r=0.05, k=20) # same parameter names in function
years = seq(from=1, to=50)

#use solver and growth rate function
results=ode(initial_pop, years, dexpopk, params_rk)

# column names
colnames(results)=c("year","P")

# save results in a dataframe

results_df <- as.data.frame(results) # same result as propose as rachel

```
<br>

With the selected parameters, the population reachs a value 7.576973 in the year 50. <br>

```{r}
# plot
ggplot(results_df, aes(year,P))+
  geom_point()+
  labs(title="Population for the next 50 years", 
       y="Population", 
       x="Years")+
  theme_bw()
```

